---
title: "Grade Analysis"
author: "Sarah Grant"
format: html
---
### Question 1
```{r}
#| message: false
here::i_am("Grade Analysis.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

### Question 2
```{r}
grades<- read.csv("grades.csv")
```
Below I am computing the minimum, maximum, median and mean grade at the Exam and including the results as a table in the rendered document.
```{r}
grade_stats <- grades %>%
  filter(!is.na(Exam)) %>%
  summarise(mingrade = min(Exam),
            maxgrade = max(Exam),
            mediangrade = median(Exam),
            meangrade = mean(Exam))

table_grade_stats<-
  knitr::kable(grade_stats,
  caption = "Grade Statistics",
  col.names = c("Minimum Grade", "Maximum Grade", "Median Grade", "Mean Grade"))

table_grade_stats
```
### Question 3

```{r}
Non_examined <- grades %>%
  filter(is.na(Exam)) %>%
  nrow()
```

The number of students who did not take the final exam is `r Non_examined`.

### Question 4
Below is the distribution of Exam grades, within the group of those who took the exam:
```{r}
Exam_grade_dist <- grades %>%
  ggplot(aes(y=Exam), filter(!is.na(Exam))) +
  geom_histogram(binwidth = 5, fill="skyblue", color= "black",) +
  labs( title= "Exam Grade Distribution", y="Grade", x= "Frequency") +
  coord_flip()

Exam_grade_dist
```
### Question 5

Below is a table depicting the number of students in each group:
```{r}
bygroup <- grades %>%
  group_by(Group) %>%
  summarise(N=n()) %>%
  arrange(N)

  knitr::kable(bygroup, caption = "Number of Students in Each Group")
```
### Question 6

Below is the same information, just represented graphically:
```{r}
bygroupgraph <- ggplot(bygroup, aes(x= Group, y = N)) +
  geom_bar(stat = "Identity", fill = "blue") +
  labs(title = "Number of Students in Each Group", x= "Group", y = "Count") +
  theme_minimal()

bygroupgraph
```


