---
title: "Grade Analysis"
author: "Sarah Grant"
format: html
---
### Question 1
```{r}
#| message: false
here::i_am("Grade Analysis.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(stringr)
```

### Question 2
```{r}
grades<- read.csv("grades.csv")
```
Below I am computing the minimum, maximum, median and mean grade at the Exam and including the results as a table in the rendered document.
```{r}
grade_stats <- grades %>%
  filter(!is.na(Exam)) %>%
  summarise(mingrade = min(Exam),
            maxgrade = max(Exam),
            mediangrade = median(Exam),
            meangrade = mean(Exam))

table_grade_stats<-
  knitr::kable(grade_stats,
  caption = "Grade Statistics",
  col.names = c("Minimum Grade", "Maximum Grade", "Median Grade", "Mean Grade"))

table_grade_stats
```
### Question 3

```{r}
Non_examined <- grades %>%
  filter(is.na(Exam)) %>%
  nrow()
```

The number of students who did not take the final exam is `r Non_examined`.

### Question 4
Below is the distribution of Exam grades, within the group of those who took the exam:
```{r}
Exam_grade_dist <- grades %>%
  ggplot(aes(y=Exam), filter(!is.na(Exam))) +
  geom_histogram(binwidth = 5, fill="skyblue", color= "black",) +
  labs( title= "Exam Grade Distribution", y="Grade", x= "Frequency") +
  coord_flip()

Exam_grade_dist
```
### Question 5

Below is a table depicting the number of students in each group:
```{r}
bygroup <- grades %>%
  group_by(Group) %>%
  summarise(N=n()) %>%
  arrange(N)

  knitr::kable(bygroup, caption = "Number of Students in Each Group")
```
### Question 6

Below is the same information, just represented graphically:
```{r}
bygroupgraph <- ggplot(bygroup, aes(x= Group, y = N)) +
  geom_bar(stat = "Identity", fill = "blue") +
  labs(title = "Number of Students in Each Group", x= "Group", y = "Count") +
  theme_minimal()

bygroupgraph
```
### Question 7 

Below is two different graphical representations of the distribution of the grades at the Exam conditioned by the group.

Firstly I am using grouped histograms (only using observations of those who took the exam):

```{r}
examdistbygroup <- grades %>%
  ggplot(aes(x= Exam, fill = Group), filter(!is.na(Exam))) +
  geom_histogram(binwidth = 1, position= "dodge", fill="skyblue", color= "black",) +
  labs(title = "Distribution of Exam Grades by Group (Grouped Histograms)", x = "Exam Grade", y= "Frequency") +
  theme_minimal()

examdistbygroup
```
Secondly, I am creating a box plot as another way to display this distribution:
```{r}
examdistboxplot <- grades %>%
  ggplot(aes(x = Group, y= Exam, fill = Group), filter(!is.na(Exam))) +
  geom_boxplot() +
  labs(title = "Distribution of Exam Grades by Group (Box Plot)", x= "Group", y= "Exam Grade") +
  theme_minimal()

examdistboxplot
```
### Question 8

Below is the computation of the number of students who missed the exam in each group, arranged in ascending order of the number of students who missed an exam in each group :

```{r}
exammissedno <- grades %>%
  group_by (Group) %>%
  summarise(N = sum(is.na(Exam))) %>%
  arrange(N)

knitr::kable(exammissedno, caption = "Number of Students Who Missed the Exam in Each Group")
```
### Question 9

Below, I am representating graphically, the table from question 8:
```{r}
exammissednograph <- ggplot(exammissedno, aes(x= Group, y= N)) +
  geom_col(fill = "blue", color ="black") +
  labs(title = "Number of Students Who Missed the Exam by Group", x = "Group", y = "Count") +
  theme_minimal()

exammissednograph
```
### Question 10

Below I am creating a data frame built from the grades data set reshaped to a long format:
```{r}
longgrades <- grades %>%
  pivot_longer(cols = c(matches("MCQ"), Exam),
               names_to = "name",
               values_to = "value")
```

### Question 11

Below is a computation of the number of missing grades in total for each student arranged in ascending order of Student ID. Since there is a huge number of students, I am only displaying the first few lines of the table on the rendered documents:
```{r}
gradesmissingbystudent <- longgrades %>%
  group_by(Id) %>%
  summarise(MissingGrades =sum(is.na(value)))

knitr::kable(head(gradesmissingbystudent))
```
### Question 12

Below I am representing graphically, the distribution of the number of missing grades per student:
```{r}
gradesmissingbystudentgraph <- gradesmissingbystudent %>%
  ggplot(aes(x = MissingGrades)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Missing Grades per Student", x = "Number of Missing Grades", y= "Frequency") +
  theme_minimal()

gradesmissingbystudentgraph
```
### Question 13

Below I am computing the number of students who missed the exam in each group using the long format as computed in question 8:
```{r}
gradesmissingbystudentstringr <- longgrades %>%
  filter(str_starts(name, "MCQ") | name== "Exam") %>%
  group_by(Group) %>%
  summarise(Count = sum(is.na(value)))

knitr::kable(gradesmissingbystudentstringr)
```
### Question 14
Below I have computed the number of missing grades for the online tests for each student, using the long format (and again, only showing the first few rows since it is a huge table):
```{r}
onlinegradesmissingbystudents <- longgrades %>%
  filter(str_starts(name, "Online_MCQ")) %>%
  group_by(Id) %>%
  summarise(MissingOnlineGrades = sum(is.na(value)))

knitr::kable(head(onlinegradesmissingbystudents))
```

