---
title: "Grade Analysis"
author: "Sarah Grant"
format: html
---
### Question 1
```{r}
#| message: false
here::i_am("Grade Analysis.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(stringr)
library(viridisLite)
```

### Question 2
```{r}
grades<- read.csv("grades.csv")
```
Below I am computing the minimum, maximum, median and mean grade at the Exam and including the results as a table in the rendered document.
```{r}
grade_stats <- grades %>%
  filter(!is.na(Exam)) %>%
  summarise(mingrade = min(Exam),
            maxgrade = max(Exam),
            mediangrade = median(Exam),
            meangrade = mean(Exam))

table_grade_stats<-
  knitr::kable(grade_stats,
  caption = "Grade Statistics",
  col.names = c("Minimum Grade", "Maximum Grade", "Median Grade", "Mean Grade"))

table_grade_stats
```
### Question 3

```{r}
Non_examined <- grades %>%
  filter(is.na(Exam)) %>%
  nrow()
```

The number of students who did not take the final exam is `r Non_examined`.

### Question 4
Below is the distribution of Exam grades, within the group of those who took the exam:
```{r}
Exam_grade_dist <- grades %>%
  ggplot(aes(y=Exam), filter(!is.na(Exam))) +
  geom_histogram(binwidth = 5, fill="skyblue", color= "black",) +
  labs( title= "Exam Grade Distribution", y="Grade", x= "Frequency") +
  coord_flip()

Exam_grade_dist
```
### Question 5

Below is a table depicting the number of students in each group:
```{r}
bygroup <- grades %>%
  group_by(Group) %>%
  summarise(N=n()) %>%
  arrange(N)

  knitr::kable(bygroup, caption = "Number of Students in Each Group")
```
### Question 6

Below is the same information, just represented graphically:
```{r}
bygroupgraph <- ggplot(bygroup, aes(x= Group, y = N)) +
  geom_bar(stat = "Identity", fill = "blue") +
  labs(title = "Number of Students in Each Group", x= "Group", y = "Count") +
  theme_minimal()

bygroupgraph
```
### Question 7 

Below is two different graphical representations of the distribution of the grades at the Exam conditioned by the group.

Firstly I am using grouped histograms (only using observations of those who took the exam):

```{r}
examdistbygroup <- grades %>%
  ggplot(aes(x= Exam, fill = Group), filter(!is.na(Exam))) +
  geom_histogram(binwidth = 1, position= "dodge", fill="skyblue", color= "black",) +
  labs(title = "Distribution of Exam Grades by Group (Grouped Histograms)", x = "Exam Grade", y= "Frequency") +
  theme_minimal()

examdistbygroup
```
Secondly, I am creating a box plot as another way to display this distribution:
```{r}
examdistboxplot <- grades %>%
  ggplot(aes(x = Group, y= Exam, fill = Group), filter(!is.na(Exam))) +
  geom_boxplot() +
  labs(title = "Distribution of Exam Grades by Group (Box Plot)", x= "Group", y= "Exam Grade") +
  theme_minimal()

examdistboxplot
```
### Question 8

Below is the computation of the number of students who missed the exam in each group, arranged in ascending order of the number of students who missed an exam in each group :

```{r}
exammissedno <- grades %>%
  group_by (Group) %>%
  summarise(N = sum(is.na(Exam))) %>%
  arrange(N)

knitr::kable(exammissedno, caption = "Number of Students Who Missed the Exam in Each Group")
```
### Question 9

Below, I am representating graphically, the table from question 8:
```{r}
exammissednograph <- ggplot(exammissedno, aes(x= Group, y= N)) +
  geom_col(fill = "blue", color ="black") +
  labs(title = "Number of Students Who Missed the Exam by Group", x = "Group", y = "Count") +
  theme_minimal()

exammissednograph
```
### Question 10

Below I am creating a data frame built from the grades data set reshaped to a long format:
```{r}
longgrades <- grades %>%
  pivot_longer(cols = c(matches("MCQ"), Exam),
               names_to = "name",
               values_to = "value")
```

### Question 11

Below is a computation of the number of missing grades in total for each student arranged in ascending order of Student ID. Since there is a huge number of students, I am only displaying the first few lines of the table on the rendered documents:
```{r}
gradesmissingbystudent <- longgrades %>%
  group_by(Id) %>%
  summarise(MissingGrades =sum(is.na(value)))

knitr::kable(head(gradesmissingbystudent))
```
### Question 12

Below I am representing graphically, the distribution of the number of missing grades per student:
```{r}
gradesmissingbystudentgraph <- gradesmissingbystudent %>%
  ggplot(aes(x = MissingGrades)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Missing Grades per Student", x = "Number of Missing Grades", y= "Frequency") +
  theme_minimal()

gradesmissingbystudentgraph
```
### Question 13

Below I am computing the number of students who missed the exam in each group using the long format as computed in question 8:
```{r}
gradesmissingbystudentstringr <- longgrades %>%
  filter(str_starts(name, "MCQ") | name== "Exam") %>%
  group_by(Group) %>%
  summarise(Count = sum(is.na(value)))

knitr::kable(gradesmissingbystudentstringr)
```
### Question 14
Below I have computed the number of missing grades for the online tests for each student, using the long format (and again, only showing the first few rows since it is a huge table):
```{r}
onlinegradesmissingbystudents <- longgrades %>%
  filter(str_starts(name, "Online_MCQ")) %>%
  group_by(Id) %>%
  summarise(MissingOnlineGrades = sum(is.na(value)))

knitr::kable(head(onlinegradesmissingbystudents))
```
### Question 15

For this question I am representing graphically the distribution of grades at the Exam conditioned by the number of missing grades for the online tests. There are two approaches to this question so I will do them both below, beginning with the dplyr row-wise functionality:
```{r}
qu15onlinegradesmissingbystudents <- grades %>%
  select(Id, starts_with("Online_MCQ")) %>%
  rowwise() %>%
  mutate(MissingOnlineGrades = sum(is.na(c_across(everything())))) %>%
  select(Id, MissingOnlineGrades)

gradeswithmissingonlineexam <- grades %>%
  left_join(qu15onlinegradesmissingbystudents, by = "Id")

qu15onlinegradesmissingbystudentsboxplot <- gradeswithmissingonlineexam %>% ggplot(aes(x = MissingOnlineGrades, y= Exam)) +
  geom_point()+
  labs(title = "Distribution of Exam Grades by Missing Online Test Grades", x= "Missing Online Test Grades", y= "Exam Grade") +
  theme_minimal()
```
The second approach is by joining the results with the original table:
```{r}

gradeswithmissingonlinetest <- grades %>%
  left_join(onlinegradesmissingbystudents, by = "Id")

gradeswithmissingonlinetestplot <- gradeswithmissingonlinetest %>%
  ggplot(aes(x =MissingOnlineGrades, y= Exam)) +
  geom_point(fill ="black") +
  labs(title = "Distribution of Exam Grades by Missing Online Test Grades", x = "Missing Online Test Grades", y="Exam Grade") + theme_minimal()

gradeswithmissingonlinetestplot
```
With both of these approaches the result is the same, so I have only included one of these graphs on the rendered document.

### Question 16

Below I am creating a table with two columns: Id to identify students, and Missed with the value "TRUE" when the student has missed at least one MCQ grade and "FALSE" if they haven't missed any MCQ grades:
```{r}
missedgradetable <- grades %>%
  select(Id, starts_with("MCQ_")) %>% 
  mutate(Missed = rowSums(is.na(.))>0) %>%
  distinct(Id, .keep_all =TRUE)%>%
  arrange(Id)
```
```{r}
missedgradetable<- select(missedgradetable, Id, Missed)

knitr::kable(head(missedgradetable))
```
The table above is very long so I am just displaying the first few columns. I have arranged the table by ID (in ascending order).

### Question 17
Below is a table with two columns: 'Group' to identify groups and 'P_missed' with the percentage of students in each group who missed at least one MCQ_xx grade. I have arranged in ascending order of the percentage of students who missed at least one MCQ grade in each group:
```{r}
group_p_missedgradetable <- missedgradetable %>%
  left_join(grades %>% select(Id, Group), by = "Id") %>%
  group_by(Group) %>%
  summarise(P_missed =mean(Missed)) %>%
  arrange(P_missed)

knitr::kable(group_p_missedgradetable)
```
### Question 18
Below is the graphical representation of the average grade at the Exam per group as a function of the percentage of missed MCQ_xx grade as defined in question 17. I added a trend line to the display to better see if there is any correlation between the average grade, and the percentage of missed MCQ grades:
```{r}
averagegradepergroup <- grades %>%
  group_by(Group) %>%
  summarise(AverageGrade = mean(Exam, na.rm = TRUE))

examvsmissedgraph <- group_p_missedgradetable %>%
  inner_join(averagegradepergroup, by = "Group") %>%
  ggplot(aes(x = P_missed*100, y= AverageGrade)) +
  geom_point() +
  geom_jitter(alpha=0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Average Exam Grade vs. Percentage of Missed MCQ Grades", x = "Percentage of Missed MCQ Grades", y = "Average Exam Grade") +
  theme_minimal()

examvsmissedgraph
```
There is a weak negative correlation (this doesn't necessarily mean that there is causation but in general, a negative relationship would be expected).
